Efficiently Handling Guards in the Low Level Design of RPython's tracing JIT

### 2012-07-30 talk/vmil2012/paper.tex
more related work

\todo{resume data size estimates on 64bit}
\todo{figure about failure counts of guards (histogram?)}
\todo{integrate high level resume data size into Figure \ref{fig:backend_data}}
\todo{add resume data sizes without sharing}
\todo{add a footnote about why guards have a threshold of 100}

@@

@@

\bivab{mention Gal et al.~\cite{Gal:2009ux} trace stitching}
and also mention \bivab{Dynamo's fragment linking~\cite{Bala:2000wv}} in
relation to the low-level guard handling.

\todo{look into tracing papers for information about guards and deoptimization}
LuaJIT \todo{link to mailing list discussion}
http://lua-users.org/lists/lua-l/2009-11/msg00089.html

>>>>

\todo{figure about failure counts of guards (histogram?)}
\todo{add resume data sizes without sharing}
\todo{add a footnote about why guards have a threshold of 100}

@@

Guards as described are a concept associated with tracing just-in-time
compilers to represent possible divergent control flow paths.

@@

Mike Pall, the author of LuaJIT describes in a post to the lua-users mailing
list different technologies and techniques used in the implementation of
LuaJIT~\cite{Pall:2009}.\todo{decide if LuaJIT is a footnote or a reference and
fix website citation} Pall explains that guards in LuaJIT use a datastucture
called snapshots, similar to PyPy's resume data, to store the information about
how to rebuild the state from a side-exit using the information in the snapshot
and the machine execution state. Pall also acknowledges that snapshot for
guards are associated with a large memory footprint. The solution used in
LuaJIT is to store sparse snapshots, avoiding the creation of snapshots for
every guard to reduce memory pressure. Snapshots are only created for guards
after updates to the global state, after control flow points from the original
program and for guards that are likely to fail. As an outlook Pall mentions the
plans to switch to compressed snapshots to further reduce redundancy.

Linking side exits to pieces of later compiled machine code was described first
in the context of Dynamo~\cite{Bala:2000wv} under the name of Fragment Linking.
Once a new hot trace is emitted into the fragment cache it is linked to side
exit that led to the compilation. Fragment Linking avoids the performance
penalty involved in leaving the compiled and it to remove the compensation
code used when restoring the machine state on a side exit.

In~\cite{Gal:2006} Gal et. al describe that in the HotpathVM they experimented
with having one generic compensation code block, like the RPython JIT, that
uses a register variable mapping to restore the interpreter state. Later this
was replaced by generting compensation code for each guard which produced a
lower overhead in their benchmarks. HotpathVM also records secondary traces
starting from failing guards that are connected directly to the original trace.
Secondary traces are compiled by first restoring the register allocator state to
the state at the side exit. The information is retrieved from a mapping stored
in the guard that maps machine level registers and stack to Java level stack
and variables.

Gal et. al~\cite{Gal:2009ux} write about how TraceMonkey uses trace stitching
to avoid th overhead of returning to the trace monitor and calling another
trace when taking a side exit. In their approach it is required to write live
values to an activation record before entering the new trace.
